---
title: "NOC language similarity vs. skill distances"
output:
  flexdashboard::flex_dashboard:
runtime: shiny
resource_files:
- out/similarity.rds
- out/tf_idf.rds
- out/dist_and_sim.rds
---

```{r}
library(tidyverse)
library(here)
library(plotly)
#functions-------------------------------
compute_pca_projection <- function(df) {
  stopifnot(
    is.data.frame(df),
    all(c("distance", "similarity") %in% names(df)),
    is.numeric(df$distance),
    is.numeric(df$similarity)
  )

  X <- df[, c("distance", "similarity")]
  pca <- prcomp(X, center = TRUE, scale. = TRUE)
  v1 <- pca$rotation[, 1]
  center <- colMeans(X)

  X_centered <- sweep(as.matrix(X), 2, center)
  proj_scalar <- as.vector(X_centered %*% v1)
  proj_coords <- proj_scalar %*% t(v1)
  proj_coords_full <- sweep(proj_coords, 2, center, "+")

  df_out <- cbind(df,
                  pca_index = proj_scalar,
                  x_proj = proj_coords_full[, 1],
                  y_proj = proj_coords_full[, 2])

  idx <- df_out$pca_index
  if (cor(df_out$similarity, idx) < 0 || cor(df_out$distance, idx) > 0) {idx <- -idx}
  df_out$pca_index <- idx
  
  list(
    data = df_out,
    pc1 = v1,
    center = center
  )
}

plot_pca_projections <- function(pca_result) {
  df <- pca_result$data
  v1 <- pca_result$pc1
  center <- pca_result$center

  plt <- ggplot(df, aes(x = distance,
                        y = similarity,
                        text = paste0("NOC: ",
                                      description2,
                                      "\n Index value: ",
                                      round(pca_index,2)))) +
    geom_point(alpha=.5) +
    geom_segment(aes(xend = x_proj, yend = y_proj), alpha = 0.1) +
    geom_point(aes(x = x_proj, y = y_proj), alpha=.5, colour="red") +
    coord_fixed() +
    theme_minimal() +
    labs(title = "PCA: Orthogonal Projections onto First Component",
         x = "Scaled Distance",
         y = "Scaled Similarity")

  ggplotly(plt, tooltip = "text")
}


#load the rds files in the out subdirectory
similarity <- read_rds(here("out", "similarity.rds"))
tf_idf <- read_rds(here("out", "tf_idf.rds"))
dist_and_sim <- read_rds(here("out", "dist_and_sim.rds"))
nocs <- unique(similarity$noc1)
descriptions <- unique(similarity$description1)
```


Language Similarity {data-orientation=rows}
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
selectInput("desc", "Select a NOC: ", choices = descriptions, selected = "Drillers and blasters - surface mining, quarrying and construction")

#reactive elements

reactive_similarity <- reactive({
   similarity|>
    filter(description1 == input$desc) |>
    slice_max(similarity, n = 10)|>
    mutate(links = paste0("https://noc.esdc.gc.ca/Structure/NOCProfile?GocTemplateCulture=en-CA&code=", noc2, "&version=2021.0"),
           links = paste0('<a href="', links, '" target="_blank">', description2, '</a>'),
           link = paste0("https://noc.esdc.gc.ca/Structure/NOCProfile?GocTemplateCulture=en-CA&code=", noc1, "&version=2021.0")
           )
})

output$dynamic_title <- renderUI({
  tags$h3(
    tags$a(href = reactive_similarity()$link[1], target = "_blank", reactive_similarity()$description1[1])
  )
})
```

### Language Similarity 

-   Top panel shows the 10 closest NOCs to the selected NOC based on language similarity.
-   Note that the [light blue](https://noc.esdc.gc.ca/) NOC descriptions are hyperlinks.
-   Bottom left panel shows the most common terms in the selected NOC.
-   Bottom middle panel shows the terms that are rare in other NOCs text.
-   Bottom right panel shows the defining terms: terms that are common in this NOC, but rare in others.

## Row

### `r uiOutput("dynamic_title")`

```{r, fig.retina=2}
renderPlotly({
  reactive_similarity()|>
    plot_ly(
      x = ~similarity,
      y = ~description2,
      type = 'bar',
      orientation = 'h',
      hoverinfo = 'none'
      )%>%
    plotly::layout(
      margin = list(t = 50),
      title = list(
      text = paste0("Top 10 similar NOCs"),
      font = list(size = 20)
      ),
      xaxis = list(title = "Cosine Similarity"),
      yaxis = list(
      title = "",
      tickvals = reactive_similarity()$description2,
      ticktext = reactive_similarity()$links,
      tickmode = "array"
      ),
      margin = list(l = 150)
      )
})

```

## Row

### Term Frequency: most common terms

```{r, fig.retina=2}
renderPlotly({
  tf_idf |>
    filter(description == input$desc) |>
    slice_max(tf, n = 10) |>
    mutate(word = fct_reorder(word, tf, .desc = TRUE)) |>
    plot_ly(
  x = ~tf,
  y = ~word,
  type = 'bar',
  orientation = 'h',
  hoverinfo = 'none'
)%>%
  plotly::layout(
    xaxis = list(title = ""),
    yaxis = list(title = "")
  )
})
```

### Inverse Document Frequency: terms that are most rare for other NOCS

```{r, fig.retina=2}
renderPlotly({
  tf_idf |>
    filter(description == input$desc) |>
    slice_max(idf, n = 10) |>
    mutate(word = fct_reorder(word, idf, .desc = TRUE)) |>
    plot_ly(
  x = ~idf,
  y = ~word,
  type = 'bar',
  orientation = 'h',
  hoverinfo = 'none'
)%>%
  plotly::layout(
    xaxis = list(title = ""),
    yaxis = list(title = "")
  )
})
```

### TF\*IDF: the defining terms

```{r, fig.retina=2}
renderPlotly({
  tf_idf |>
    filter(description == input$desc) |>
    slice_max(tf_idf, n = 10) |>
    mutate(word = fct_reorder(word, tf_idf, .desc = TRUE)) |>
    plot_ly(
  x = ~tf_idf,
  y = ~word,
  type = 'bar',
  orientation = 'h',
  hoverinfo = 'none'
)%>%
  plotly::layout(
    xaxis = list(title = ""),
    yaxis = list(title = "")
  )
})
```

Language Similarity vs Skills Distance  {data-orientation=columns}
==============================================


Inputs {.sidebar}
-------------------------------------

### Language Similarity vs Skills Distance

-  ONET skills, knowledge, abilities and work activities scores can be used to 
measure distance between NOCs.
-  Here we create an index of similarity based both on language similarity and proximity in terms of skills, knowledge, abilities, and work activities.
- The index is computed as a projection onto the first principal component of the PCA of the distance and similarity scores. (the red dots)
- Top left: similar NOCs
- Bottom right: the most different NOCs.
- Points below the diagonal are more similar in skills, but less similar in language.
- Points above the diagonal are more similar in language, but less similar in skills.

```{r}
#reactive elements------------
pca_result <- reactive({
  dist_and_sim |>
    filter(description1 == input$desc) |>
    compute_pca_projection()
})

most_similar <- reactive({
  pca_result()$data|>
    slice_max(pca_index, n = 10)
})
```



## Column {data-width=650}

###

```{r}
renderPlotly({
  plot_pca_projections(pca_result()) 
})
```

## Column

###

```{r}
renderPlot({
 most_similar() |>
  ggplot(aes(x=pca_index, y=fct_reorder(description2, pca_index))) +
    geom_col(alpha=.5)+
    theme_minimal()+
    labs(title = "Top 10 NOCs by PCA Index",
         x = "PCA Index",
         y = "")
})
```




